SEER Medicaid

### 2024-01-11 Manuscript Analysis code 0 to 19.Rmd
latest version after minor clean-up

# Open libraries
```{r}
# install.packages("pacman")
pacman::p_load(readr, dplyr, tidyverse, lubridate, survival, survminer, hablar, table1, patchwork, openxlsx, DiagrammeR, lmtest, forestplot, MASS, nnet, broom, forcats, sandwich, mlogit, sjstats, odds.n.ends, labelled, readxl)
```
# Functions used
```{r}
# function to calculate clustered SEs (from: https://stackoverflow.com/questions/73136246/clustered-standard-errors-stars-and-summary-statistics-in-modelsummary-for-mul)
mlogit.clust <- function(model,data,variable) {
  beta <- c(t(coef(model)))
  vcov <- vcov(model)
  k <- length(beta)
  n <- nrow(data)
  max_lev <- length(model$lev)
  xmat <- model.matrix(model)
  # u is deviance residuals times model.matrix
  u <- lapply(2:max_lev, function(x)
    residuals(model, type = "response")[, x] * xmat)
  u <- do.call(cbind, u)
  m <- dim(table(data[,variable]))
  u.clust <- matrix(NA, nrow = m, ncol = k)
  fc <- factor(data[,variable])
  for (i in 1:k) {
    u.clust[, i] <- tapply(u[, i], fc, sum)
  }
  cl.vcov <- vcov %*% ((m / (m - 1)) * t(u.clust) %*% (u.clust)) %*% vcov
  return(cl.vcov = cl.vcov)
}
```

@@

x <- unique(df$patient_id) # 841 vector of unique ids who had enrollment in multiple states to exclude in exclusion section
```

@@

# FIGURE 1
```{r}
Figure1<-grViz("digraph flowchart {
      # node definitions with substituted label text
      node [fontname = Helvetica, shape = rectangle, fontsize=16]        
      tab1 [label = '@@1']
      tab2 [label = '@@2']
      tab3 [label = '@@3']
      tab4 [label = '@@4']
      tab5 [label = '@@5']
      tab6 [label = '@@6']
      tab7 [label = '@@7']
      # edge definitions with the node IDs
      tab1 -> tab2 -> tab3 -> tab4 -> tab5 -> tab6 -> tab7;
      }
      [1]: 'Records received for patients diagnosed with malignant cancers at ages 0 to 19 years from 2006-2013 n = 34,196'
      [2]: 'Excluding 1800 individuals whose cancer was not their first malignant primary was n = 32,396' 
      [3]: 'Excluding 16 individuals whose cancer sequence number was not in 0 or 1 n = 32,380'
      [4]: 'Excluding 818 individuals linked to multiple state Medicaid files n = 31,562'
      [5]: 'Excluding 152 individuals with Medicaid enrollment match scores <5 n = 31,410'
      [6]: 'Excluding 163 individuals with missing data on cancer specific cause of death n = 31,247'
      [7]: 'Excluding 593 individuals individuals with missing data on Yost ACS 2006 to 2010 Quintile n = 30,654'
      ")
Figure1
```

@@

# SUPPLEMENTAEY TABLE 1:
```{r}
SuppTable1 <- rbind(SuppTable1a, SuppTable1b)

@@

# FIGURE 2:  Stage analysis by cancer type mcont66
```{r}
# graph ORs and 95% CIs
SuppTable1a$`Exposure comparison` <- factor(SuppTable1a$`Exposure comparison`, levels = c("Enrolled vs. Not enrolled", "Discontinuous vs. Not enrolled", "Other vs. Not enrolled", "Continuous vs. Not enrolled"))

@@

# the if_elsee function does not like labels
seer0to19 <- seer0to19 %>% 
  remove_labels()

>>>>

# Open libraries
```{r}
# install.packages("pacman")
# install.packages("pacman")
pacman::p_load(readr, dplyr, tidyverse, lubridate, survival, survminer, hablar, table1, patchwork, openxlsx, DiagrammeR, lmtest, forestplot, MASS, nnet, broom, forcats, sandwich, mlogit, sjstats, odds.n.ends, labelled, readxl)
```
# Functions used
```{r}
# function to calculate clustered SEs (from: https://stackoverflow.com/questions/73136246/clustered-standard-errors-stars-and-summary-statistics-in-modelsummary-for-mul)
mlogit.clust <- function(model,data,variable) {
  beta <- c(t(coef(model)))
  vcov <- vcov(model)
  k <- length(beta)
  n <- nrow(data)
  max_lev <- length(model$lev)
  xmat <- model.matrix(model)
  # u is deviance residuals times model.matrix
  u <- lapply(2:max_lev, function(x)
    residuals(model, type = "response")[, x] * xmat)
  u <- do.call(cbind, u)
  m <- dim(table(data[,variable]))
  u.clust <- matrix(NA, nrow = m, ncol = k)
  fc <- factor(data[,variable])
  for (i in 1:k) {
    u.clust[, i] <- tapply(u[, i], fc, sum)
  }
  cl.vcov <- vcov %*% ((m / (m - 1)) * t(u.clust) %*% (u.clust)) %*% vcov
  return(cl.vcov = cl.vcov)
}
```

@@

x <- unique(df$patient_id) # 3588 vector of unique ids who had enrollment in multiple states to exclude in exclusion section
```

@@

# FIGURE 1
```{r}
Figure1 <- grViz("digraph flowchart {
      # node definitions with substituted label text
      node [fontname = Helvetica, shape = rectangle, fontsize=16]        
      tab1 [label = '@@1']
      tab2 [label = '@@2']
      tab3 [label = '@@3']
      tab4 [label = '@@4']
      tab5 [label = '@@5']
      tab6 [label = '@@6']
      tab7 [label = '@@7']
      # edge definitions with the node IDs
      tab1 -> tab2 -> tab3 -> tab4 -> tab5 -> tab6 -> tab7;
      }
      [1]: 'Records received for patients diagnosed with malignant cancers at ages 0 to 19 years from 2006-2013 n = 34,196'
      [2]: 'Excluding 1800 individuals whose cancer was not their first malignant primary was n = 32,396' 
      [3]: 'Excluding 16 individuals whose cancer sequence number was not in 0 or 1 n = 32,380'
      [4]: 'Excluding 818 individuals linked to multiple state Medicaid files n = 31,562'
      [5]: 'Excluding 152 individuals with Medicaid enrollment match scores <5 n = 31,410'
      [6]: 'Excluding 163 individuals with missing data on cancer specific cause of death n = 31,247'
      [7]: 'Excluding 593 individuals individuals with missing data on Yost ACS 2006 to 2010 Quintile n = 30,654'
      ")

Figure1

library(DiagrammeRsvg)  # for conversion to svg
library(rsvg)  # for saving svg

Figure1 %>%
   DiagrammeRsvg::export_svg() %>% 
  charToRaw() %>% 
  rsvg::rsvg_pdf("Figure1.pdf")
```

@@

# SUPPLEMENTAEY TABLE 1: Stage at diagnosis cancer type analysis
```{r}
SuppTable1 <- rbind(SuppTable1a, SuppTable1b)

@@

# FIGURE 2:  Stage analysis by cancer type mcont66
```{r}
# graph ORs and 95% CIs
SuppTable1$`Exposure comparison` <- factor(SuppTable1$`Exposure comparison`, levels = c("Enrolled vs. Not enrolled", "Discontinuous vs. Not enrolled", "Other vs. Not enrolled", "Continuous vs. Not enrolled"))

@@

# the if_else function does not like labels
seer0to19 <- seer0to19 %>% 
  remove_labels()
